---
title: "Covid91 vaccine study Final2025"
author: "Issa Darrouj"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Coding details

Separating the variables into the four groups they belonged to was crucial for an effective evaluation of the data. "M", "F", "gay", and "drug users" are the four variables I created. To determine whether or not they should get the Covid-19 vaccine, they were kept apart. I did this in order to generate the efficacy intervals. They were kept in different variables. 

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

### Barplot

```{r}
dd2 <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot displays the male population and the correlation between contracting COVID-19 and receiving a placebo or the vaccination. 

The first graph displays the number of individuals who either had COVID-19 or were well, as well as whether or not they received a placebo. The second graph displays the proportion of individuals who received either the COVID-19 vaccination or a placebo, as well as whether or not they had COVID. They are divided into groups based on whether they received the Covid-19 vaccination or a placebo. Of those who contracted COVID-19, those who received a placebo were plotted as a large bar, at around 75%, while those who received the vaccination were much smaller, at about 25%. There was a 50/50 divide between those who received the vaccination and those who took a placebo among those who were okay. 

I used a lot of codes to plot this. The code block that starts with dd2, which appears to be where the environment pane is used and contains four variables. Treatment, infected, count, and percent (which stands for percent) are some of these factors. The basicC is the name of the second group. It has a list of eleven variables that may be seen in the environment section. This is to establish the initial plan. The final set of codes is the basicCC, which is likewise located in the files pane. It has eleven variables that are useful for making the second plot, the percentage plot. 


### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```

The exact percentages of guys who contracted COVID-19 after receiving either the vaccination or a placebo are displayed in these numerical results. 

Compared to the percentage of individuals who contracted COVID-19 after receiving the vaccination (26.20%), the number of men who contracted the virus after receiving a placebo was a staggering 73.80%. Of the guys who were well, 49.99% of those who received the vaccination and 50.01% of those who received a placebo were well. Given that only 26.20% of those who received the vaccination contracted COVID-19, this indicates that the vaccine is successful in men. COVID-19 was far more common in those who took the placebo. This is crucial. 

### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Msfish <- fisher.test(table1)
Msfish
```

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions